### 2012-11-29 paper.tex
Intro skeleton

\documentclass{article}

\begin{document}

\title{Implementation Design Patters for Pursuit-Evasion Problems with Bounded Speed}
\author{Daniel Savage\\CSC 426\\{V00701453}}
\date{November 2012}
\maketitle
\pagebreak

\section{Introduction}
-- Explanation of pursuit-evasion problem in general.
-- The two various methods for analyzing the problem
-- a solution for blending the two solutions together

\section{Paradigm 1 - Discrete Formulation}

\section{Paradigm 2 - Formulation on the Continuous Plane}
Hello World!

\section{Blended Paradigm Concept}
Hello World!

\section{Conclusion}
Hello World!

\end{document}