### 2012-11-29 paper.tex
filled out skeleton
(woo)

\documentclass{article}

\begin{document}

\title{Implementation Design Patters for Pursuit-Evasion Problems with Bounded Speed}
\author{Daniel Savage\\CSC 426\\{V00701453}}
\date{November 2012}
\maketitle
\pagebreak

\section{Introduction}
% Explanation of pursuit-evasion problem in general.
% The two various methods for analyzing the problem
% a solution for blending the two solutions together

\section{Paradigm 1 - Discrete Formulation}
% explanation of the paradigm
% mathematical definitions and/or data structures
% general rules of "play"
% planar graphs need max three pursuers

\section{Paradigm 2 - Formulation on the Continuous Plane}
% explanation of the paradigm
% mathematical definitions and/or data structures
% general setup and rules of "play"
% visibility polygons and gap management
% area contamination and management
% NP-hardness

\section{Blended Paradigm Concept}
% explanation of the paradigm (NP-hardness aleviation)
% mathematical definitions and/or data structures
% general setup and rules of "play"
% delunay triangulation and minimum spanning tree connection
% visibility and contamination representation
% tesslation through barycenters
% strategic issues

\section{Conclusion}
% no idea what you're goona write here

\section{References}

\end{document}